\documentclass[12pt]{report}
\usepackage{blindtext}
\usepackage{hyperref}
\usepackage{tocloft}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage[newfloat]{minted}
\usepackage{caption}
\usepackage{booktabs}
\usepackage[toc, page]{appendix}
\usepackage{multirow}
\newenvironment{longlisting}{\captionsetup{type=listing}}{}
\title{%
  GPT-2 Based Story Teller\\
  \large CTEC3451 Development Project \\~\\
    De Montfort University} 
\author{Artem Bobrov (P2547788)}
\date{May 5 2022}
\setlength{\parindent}{1em}
\begin{document}
\maketitle
\thispagestyle{empty}
\clearpage
\tableofcontents
\setcounter{tocdepth}{1}
\thispagestyle{empty}
\clearpage
\section*{Abstract}
\addcontentsline{toc}{section}{Abstract}
\paragraph{}
The project described in this report is a story-telling application that uses the GPT-2 model to generate text. 
The model is trained on multiple datasets and is able to generate text from a variety of topics in the backend. 
The frontend is a web application that allows users to endlessly generate stories accompanied by thematic pictures.
The point of the project is to prove the concept that neural networks, if trained properly, are able to generate 
data that is very similar to human language.

\section*{Introduction}
\addcontentsline{toc}{section}{Introduction}
\paragraph{}
Neural networks weren't so popular in the early days of computing, but they are now a common tool for machine learning.
First neural network was invented by Frank Rosenblatt in the late 1950s \citep{rosenblatt_1958_the}. Perceptron has only 
one layer of neurons, but more advance neural networks can have multiple layers. Before Transformer neural networks 
there were many different types of neural networks (e.g. Convolutional, Recurrent, LSTM, Feed Forward, etc).

Transformers were introduced in 2017 by Google \citep{attention_is_all_you_need} and are a new type of neural network. It uses self-attention mechanism 
to focus on the most relevant information in the input. It weights the significance of each part of the input differently.

This report focuses on the GPT-2 model, which is a transformer type neural network developed by OpenAI in 2019 \citep{radford_wu_child_luan_amodei_sutskever_2019}.
Despite the fact that GPT-3 is a newer model, it's predecessor is still a popular choice for text generation.

An application will be developed that utilises the features of GPT-2 and wraps it up in a user-friendly web
interface that allows users to read freshly generated stories accompanied by thematic pictures. The development process 
will be examined and the results will be presented and discussed in depth. The usage of the key elements and techniques used both in
backend and frontend will be explained. Moreover the project development and research cycle will be reflected on and 
other approaches that possibly could have been used to achieve different or similar results will be presented.

\clearpage

\section*{Objectives and Goals}
\addcontentsline{toc}{section}{Objectives and Goals}

\subsection*{Initial approach}
\addcontentsline{toc}{subsection}{Initial approach}
\paragraph{}
It is important to state that the original plan for the project was to develop a similar application that generates
pictures with attached text to them (known as `memes') and allows users to watch them using a web interface. It has been
planned to use different approach to build a repository of training data using `Optical Character Recognition (OCR)' \citep{tesseract_article} and 
`Pillow' library to attach processed text to the pictures. It has been decided to adjust the original plan and change the
topic from memes to stories.

\subsection*{New approach}
\addcontentsline{toc}{subsection}{New approach}
\paragraph{}
The new approach doesn't require any massive changes to the original plan and the project will be developed using the same
techniques - indetical frontend and almost the same backend. The only difference is that OCR will be removed from the backend
because it's not needed for the project anymore. To minimize the deviation from the original plan, there will be an
additional section in the web interface that creates `memes' based on novels and other piece of art.


\subsection*{Reasons}
\addcontentsline{toc}{subsection}{Reasons}
\paragraph{}
Despite the fact that many of the features used in the initial approach have been already developed - Tesseract OCR, 
Pillow Image placing, `Meme' scraper - one major issue made it extremely perplexing to continue the project without changing the approach.

The first puzzling part was hidden in reading the text (using Tesseract OCR) from the scraped images and translate them to text strings.
Memes contain text captions that makes it easy for a human to read, but this same process is hard for machines to do so - 
especially via OCR. There were some images successfully processed by it but it didn't satisfy the requirement because the the success rate was too low.
Many solutions were tried, but even pre-processing the images before feeding them to the OCR engine didn't seem to give much improvement.
The second complexity was to build a meme repository itself. Using the meme scraper is not the best solution because
the content obtained by it can be anything. Unfiltered, uncensored - it's a huge ethical problem.

\subsection*{General Description}
\addcontentsline{toc}{subsection}{General Description}
\paragraph{}
Two general components of the development project are described in two sections. Such apparent separation between these categories 
is due to the fact that this approach allows a simpler, faster and more effiecient implementation with better user
experience, it also improves code readability and maintainability.

The first one is the backend, which is responsible for generating text based on given datasets 
(stories, novels, other pieces of art, etc). It is implemented in Python and uses numerous libraries
to achieve the goal. Multiple steps are involved in generating simple text samples including hours of training.

The second component is the frontend, which is a web application that creates a user-friendly, simple interface
to scroll and read stories. It is implemented in JavaScript using many libraries, as well as the Python part.
Such approach to frontend was chosen because web development is flexible and it's easy to adapt to different
platforms. Also it's possible to extend it to a web service that doesn't require downloading the application
to user's computer.

The goal is create a system with pre-trained model, which the user can use to scroll through the stories and 
react to each story leaving a positive or a negative feedback - using the `Like' and `Dislike' buttons. Afterwards,
these reactions will be stored in a database to be used by the developer to understand which stories should be
corrected and improved. The feedback system is designed to create a reflection on the pre-trained model as a whole and
not on separate elements (stories). Given the feedback, the developer can decide to improve the model or not.

\clearpage

\section*{Major Components}
\addcontentsline{toc}{section}{Major Components}

\subsection*{Backend}
\addcontentsline{toc}{subsection}{Backend}

\subsubsection*{GPT-2}
\addcontentsline{toc}{subsubsection}{GPT-2}
\paragraph{}
% Libraries used to operate with the model
GPT-2 is the version of the GPT family used in the project. To operate with the network it was chosen to use a 
library called `GPT-2-simple' \citep{gpt-2-simple-git} since it features an easy approach to control and train the model.
The library is written in Python and offers a documentation which was used in order to work with the neural network.
GPT-2-simple offers two types of models - one with 124 million parameters and one with 324 million parameters. 
Acknowledging the computational limits of the processing unit used - graphical (GPU), the first model with 124 million
parameters was used. It is possible to use the second model but it requires a GPU with significantly more processing power and
memory. 

It is possible to use techniques that allow to save memory and train bigger models using memory saving gradients -
the main memory loss in the computation is due to calculating the gradient of the loss by backpropagation. At the
cost of one additional forward pass, the memory consumption can be reduced significantly \citep{chen_2016_training}.
However Tensorflow framework (version 2), which is used by GPT-2-simple, is not optimized for such tasks. 

\subsubsection*{Datasets}
\addcontentsline{toc}{subsubsection}{Datasets}
\paragraph{}
% Alice, Catcher, Iliad


\subsubsection*{Pillow}
\addcontentsline{toc}{subsubsection}{Pillow}
\paragraph{}
% Meme section

\subsection*{Frontend}
\addcontentsline{toc}{subsection}{Frontend}

\subsubsection*{Web Application appearence}
\addcontentsline{toc}{subsubsection}{Web Application appearence}
\paragraph{}
% Bootstrap, Electron, theme change
The application appearance is based on Bootstrap 5 and Electron. Bootstrap is a CSS framework that is used to create
a responsive and mobile-first design. The main navigation bar that is positioned at the top of the page and can be
seen at all times since its fixed is designed using Bootstrap classes.


\subsubsection*{Displaying the content}
\addcontentsline{toc}{subsubsection}{Displaying the content}
\paragraph{}
% Lazy loading, formatting the NN output, JQuery

\subsubsection*{User feedback}
\addcontentsline{toc}{subsubsection}{User feedback}
\paragraph{}
% Like and dislike buttons

\subsubsection*{Database}
\addcontentsline{toc}{subsubsection}{Database}
\paragraph{}
% Storing settings, user feedback, theme change, JSON

\subsection*{Linking the frontend and backend}
\addcontentsline{toc}{subsection}{Linking the frontend and backend}
\paragraph{}
% Eel, Python, JavaScript





% IMPLEMENTATION:
% Building repository
% Training the model
% Generating text (best samples)
% Testing
% 

%TC:ignore

\addcontentsline{toc}{section}{Bibliography}
\bibliography{reference}{}
\bibliographystyle{agsm}

\begin{appendices}
\section*{Appendix A}
\label{appendix:A}
\addcontentsline{toc}{section}{Appendix A}

\clearpage

\section*{Appendix B}
\label{appendix:B}
\addcontentsline{toc}{section}{Appendix B}

\end{appendices}

%TC:endignore

\end{document}